\documentclass[12pt,preprint,dvipdf]{aastex}
\newcounter{address}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\units}[1]{\mathrm{#1}}
\renewcommand{\mag}{\units{mag}}
\renewcommand{\arcmin}{\units{arcmin}}
\newcommand{\rfifty}{r_{50}}
\newcommand{\rninety}{r_{90}}
\newcommand{\conc}{C}

\begin{document}

\title{
       The SDSS Large Galaxy Atlas
      }
\author{
        Ekta Patel\altaffilmark{\ref{CCPP}},
        David Mykytyn\altaffilmark{\ref{CCPP}},
        David W. Hogg\altaffilmark{\ref{CCPP},\ref{MPIA},\ref{email}},
        Dustin Lang\altaffilmark{\ref{CMU}}
       }
\setcounter{address}{1}
\altaffiltext{\theaddress}{\stepcounter{address}\label{CCPP} Center
  for Cosmology and Particle Physics, Department of Physics, New York
  University, 4 Washington Place, New York, NY 10003}
\altaffiltext{\theaddress}{\stepcounter{address}\label{MPIA}
  Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117
  Heidelberg, Germany}
\altaffiltext{\theaddress}{\stepcounter{address}\label{email} To whom
  correspondence should be addressed: \texttt{david.hogg@nyu.edu}}
\altaffiltext{\theaddress}{\stepcounter{address}\label{CMU} Carnegie
  Mellon University}

\begin{abstract}
We present the \project{SDSS Large Galaxy Atlas}, which contains
accurate positions and photometry for galaxies with half-light radii
greater than $1~\arcmin$.
\end{abstract}

\section{Introduction}

... The \project{SDSS} never made good photometry for the brightest galaxies ... The NS Atlas ...

... The \project{SDSS} made beautiful images ...

... Related work:  2MASS ... anything else? ...

The \textit{GALEX} Ultraviolet Atlas of Nearby Galaxies provides photometry, surface brightness and color profiles for 1034 nearby galaxies with a diameter larger than 1' and a surface brightness in the b-band, $u_B$=25 mag arcsec$^{-2}$ . The galaxies chosen for this data set come from \textit{GALEX} satellite, which observed galaxies in the far-ultra violet(FUV) and near-ultra violet(NUV) bands. Supplemental galaxies were chosen from other \textit{GALEX} initiatives that viewed the same fields of the sky with equivalent depth. All galaxies with a diamater larger than 1' from the Third Reference Catalog of Bright Galaxies (RC3) were also added to the sample. A total of 81 out of 1136 galaxies were not included in the Atlas because they were in regions of very high background, high Galactic extinction, extremely low UV surface brightness, or simply had images that were not of sufficient quality for analysis. 

...positions,axis diameters,morphological type, distance...
...getting magnitudes...
...galaxy type and shape...
...measuring surface brightness...
...properties by type...
...color-mag and color-color plots?...
...dust extinction...


\section{Data and input catalogs}

... Hogg: Stuff about the \project{SDSS} imaging ...

... Mykytyn: How did you make the input catalog(s)? ...

The input catalog was formed out of the union of the Third Reference Catalog of Bright Galaxies (RC3), the NASA-Sloan Atlas (NSA), and the 2MASS Large Galaxy Atlas (LGA). Objects were searched for in those catalogs with an angular diameter greater than or equal to two arcminutes. For the Third Reference Catalog the cut made was based on the measured apparent major isophotal diameter of the surface brightness $\mu_{B}$ greater than 2 arcminutes. For the NASA-Sloan Atlas the cut made was based on the 50\% light radius of the SERSIC fit. Finally for the 2MASS Large Galaxy Atlas the cut was made based on the isophotal fiduciary radius in the K-band, again cutting to a radius of 1 arcminute.

\section{Method}

... Refer to the method paper but summarize highlights ...

... Output vetting and re-running and manual tweaking---what did we do and why? ...

... A few examples showing what works and what doesn't ...

\section{Catalog}

... The big table ...

... Also exceptions tables ...

\section{Discussion}

\end{document}
